Reconstructing the Lensing Mass in the Universe \\ from Photometric Catalog Data

### 2012-11-12 doc/paper-lcr.tex
replace accidentally deleted files

\section{Theoretical Background}
\label{sec:theory}

Convergence, or Ricci focusing, occurs when a gravitational lens focuses
the rays in a given bundle. This focusing can cause distant objects to
appear brighter, and larger than they would if the lens were removed.
The convergence from an isolated mass sheet at a position
$\bmath{\theta}$ on the sky is the ratio of the projected surface
mass density ($\Sigma(D_{\mathrm{od}}\bmath{\theta})$) \devided by the
critical surface mass density at the redshift of the mass sheet
($\Sigma_{\rm cr}$),
\be
\kappa(\bmath{\theta})= \frac {\Sigma(D_{\mathrm{od}}\bmath{\theta})}
                              {\Sigma_{\rm cr}(\zd,\zs)}
\ee
where 
\be 
\label{eq:sigcrit} 
\Sigma_{\rm cr}(\zd,\zs) \equiv \frac{c^2 D_{\rm os}}{4 \pi G D_{\rm od} D_{\rm ds}}.
\ee
and the $D$'s are the angular diameter distances between the objects
refered to in the subscripts: o is the observer, d is the deflector and
s is the source. If the surface mass density of the sheet exceeds
$\Sigma_{\rm cr}$, multiple images of the source can occur, otherwise
only one image will be observed, although that image will still be
perturbed relative to the unlensed case.

\comments{
Strong lensing occurs when a massive object and background source are almost perfectly
aligned along a line of sight. The light from the background source is deflected by the
lens galaxy; this deflection allows multiple images of the background source to form
at stationary points of the time delay function. For an isolated lens, the time delay
function can be calculated from
\be \label{eq:T} 
\Delta t(\bmath{\theta},\bmath{\beta}) = \frac {1}{c} \frac{D_{\rm od} D_{\rm os}}{D_{\rm ds}} (1+z_{\rm d})\, \phi(\bmath{\theta},\bmath{\beta}),
\ee
where $\bmath{\theta}$ is the observed source position, $\bmath{\beta}$ is the 
unlensed source position, $z_{\rm d}$ is the redshift of the lens, $\phi(\bmath{\theta},\bmath{\beta})$ is
the Fermat potential. The Fermat potential is given by
\be \label{eq:FP}
\phi(\bmath{\theta},\bmath{\beta})\equiv \left[\frac{(\bmath{\theta}-\bmath{\beta})^2}{2}-\psi(\bmath{\theta}) \right], 
\ee
where $\psi(\bmath{\theta})$ is the lens potential, derived from the projected dimensionless
surface mass density, $\kappa(\bmath{\theta})$, by 
\be \label{eq:psikappa}
\kappa(\bmath{\theta})=\frac{1}{2}\nabla^2\psi(\bmath{\theta}).
\ee
}

Strong lenses, type Ia supernovae and other cosmologically interesting objects
are never viewed unlensed: 
there is some mass along the line of sight to them coming from intervening
structures. These structures may be physically associated with the target
e.g. other members of a group or cluster, or they may be the result of
chance alignments at any redshift.

\citep{Vale+White2003,HilbertEtal2007}. These additional structures
induce an external shear which can be recovered in the mass-modelling of
a strong lens, but they also introduce a convergence that is undetectable even
in these systems:
the lensed image positions, shapes and relative fluxes are affected, but not
in an unambiguous way
\citep{FalcoEtal1985}.

>>>>

\section{Theoretical Background}
\label{sec:theory}
\comments{
Convergence, or Ricci focusing, occurs when a gravitational lens focuses
the rays in a given bundle. This focusing can cause distant objects to
appear brighter, and larger than they would if the lens were removed.
The convergence from an isolated mass sheet is the ratio of the projected surface
mass density ($\Sigma$) \devided by the
critical surface mass density at the redshift of the mass sheet
($\Sigma_{\rm cr}$),
\be
\kappax= \frac {\Sigma}
                              {\Sigma_{\rm cr}(\zd,\zs)}
\ee
where 
\be 
\label{eq:sigcrit} 
\Sigma_{\rm cr}(\zd,\zs) \equiv \frac{c^2 D_{\rm os}}{4 \pi G D_{\rm od} D_{\rm ds}}.
\ee
and the $D$'s are the angular diameter distances between the objects
refered to in the subscripts: o is the observer, d is the deflector and
s is the source. If the surface mass density of the sheet exceeds
$\Sigma_{\rm cr}$, multiple images of the source can occur, otherwise
only one image will be observed, although that image will still be
perturbed relative to the unlensed case.

Strong lensing occurs when a massive object and background source are almost perfectly
aligned along a line of sight. The light from the background source is deflected by the
lens galaxy; this deflection allows multiple images of the background source to form
at stationary points of the time delay function. For an isolated lens, the time delay
function can be calculated from
\be \label{eq:T} 
\Delta t(\bmath{\theta},\bmath{\beta}) = \frac {1}{c} \frac{D_{\rm od} D_{\rm os}}{D_{\rm ds}} (1+z_{\rm d})\, \phi(\bmath{\theta},\bmath{\beta}),
\ee
where $\bmath{\theta}$ is the observed source position, $\bmath{\beta}$ is the 
unlensed source position, $z_{\rm d}$ is the redshift of the lens, $\phi(\bmath{\theta},\bmath{\beta})$ is
the Fermat potential. The Fermat potential is given by
\be \label{eq:FP}
\phi(\bmath{\theta},\bmath{\beta})\equiv \left[\frac{(\bmath{\theta}-\bmath{\beta})^2}{2}-\psi(\bmath{\theta}) \right], 
\ee
where $\psi(\bmath{\theta})$ is the lens potential, derived from the projected dimensionless
surface mass density, $\kappa(\bmath{\theta})$, by 
\be \label{eq:psikappa}
\kappa(\bmath{\theta})=\frac{1}{2}\nabla^2\psi(\bmath{\theta}).
\ee
}

External convergence is undetectable even
in strong lens systems:
the lensed image positions, shapes and relative fluxes are affected, but not
in an unambiguous way
\citep{FalcoEtal1985}.
